# Many models

## Introduction

In this chapter you're going to learn three powerful ideas that help you to work with large numbers of models with ease:

1.  Using many simple models to better understand complex datasets.

1.  Using list-columns to store arbitrary data structures in a data frame.
    For example, this will allow you to have a column that contains linear 
    models.
   
1.  Using the __broom__ package, by David Robinson, to turn models into tidy 
    data. This is a powerful technique for working with large numbers of models
    because once you have tidy data, you can apply all of the techniques that 
    you've learned about earlier in the book.

We'll start by diving into a motivating example using data about life expectancy around the world. It's a small dataset but it illustrates how important modelling can be for improving your visualisations. We'll use a large number of simple models to partition out some of the strongest signals so we can see the subtler signals that remain. We'll also see how model summaries can help us pick out outliers and unusual trends.

The following sections will dive into more detail about the individual techniques:

1. In [list-columns], you'll learn more about the list-column data structure,
   and why it's valid to put lists in data frames.
   
1. In [creating list-columns], you'll learn the three main ways in which you'll
   create list-columns.
   
1. In [simplifying list-columns] you'll learn how to convert list-columns back
   to regular atomic vectors (or sets of atomic vectors) so you can work
   with them more easily.
   
1. In [making tidy data with broom], you'll learn about the full set of tools
   provided by broom, and see how they can be applied to other types of 
   data structure.

This chapter is somewhat aspirational: if this book is your first introduction to R, this chapter is likely to be a struggle. It requires you have to deeply internalised ideas about modelling, data structures, and iteration. So don't worry if you don't get it --- just put this chapter aside for a few months, and come back when you want to stretch your brain. 

### Prerequisites

Working with many models requires many of the packages of the tidyverse (for data exploration, wrangling, and programming) and modelr to facilitate modelling.

```{r setup, message = FALSE}
library(modelr)
library(tidyverse)
```

## List-columns

Now that you've seen a basic workflow for managing many models, let's dive back into some of the details. In this section, we'll explore the list-column data structure in a little more detail. It's only recently that I've really appreciated the idea of the list-column. List-columns are implicit in the definition of the data frame: a data frame is a named list of equal length vectors. A list is a vector, so it's always been legitimate to use a list as a column of a data frame. However, base R doesn't make it easy to create list-columns, and `data.frame()` treats a list as a list of columns:.

```{r}
data.frame(x = list(1:3, 3:5))
```

You can prevent `data.frame()` from doing this with `I()`, but the result doesn't print particularly well:

```{r}
data.frame(
  x = I(list(1:3, 3:5)), 
  y = c("1, 2", "3, 4, 5")
)
```

Tibble alleviates this problem by being lazier (`tibble()` doesn't modify its inputs) and by providing a better print method:

```{r}
tibble(
  x = list(1:3, 3:5), 
  y = c("1, 2", "3, 4, 5")
)
```

It's even easier with `tribble()` as it can automatically work out that you need a list:

```{r}
tribble(
   ~x, ~y,
  1:3, "1, 2",
  3:5, "3, 4, 5"
)
```

List-columns are often most useful as intermediate data structure. They're hard to work with directly, because most R functions work with atomic vectors or data frames, but the advantage of keeping related items together in a data frame is worth a little hassle.

Generally there are three parts of an effective list-column pipeline:

1.  You create the list-column using one of `nest()`, `summarise()` + `list()`,
    or `mutate()` + a map function, as described in [Creating list-columns].

1.  You create other intermediate list-columns by transforming existing
    list columns with `map()`, `map2()` or `pmap()`. For example, 
    in the case study above, we created a list-column of models by transforming
    a list-column of data frames.
    
1.  You simplify the list-column back down to a data frame or atomic vector,
    as described in [Simplifying list-columns].

## Creating list-columns

Typically, you won't create list-columns with `tibble()`. Instead, you'll create them from regular columns, using one of three methods: 

1.  With `tidyr::nest()` to convert a grouped data frame into a nested data 
    frame where you have list-column of data frames.
    
1.  With `mutate()` and vectorised functions that return a list.

1.  With `summarise()` and summary functions that return multiple results. 

Alternatively, you might create them from a named list, using `tibble::enframe()`.

Generally, when creating list-columns, you should make sure they're homogeneous: each element should contain the same type of thing. There are no checks to make sure this is true, but if you use purrr and remember what you've learned about type-stable functions, you should find it happens naturally.

### With nesting

`nest()` creates a nested data frame, which is a data frame with a list-column of data frames. In a nested data frame each row is a meta-observation: the other columns give variables that define the observation, and the list-column of data frames gives the individual observations that make up the meta-observation.

There are two ways to use `nest()`. When applied to a grouped data frame, `nest()` keeps the grouping columns as is, and bundles everything else into the list-column.

You can also use it on an ungrouped data frame, specifying which columns you want to nest.

### From vectorised functions

Some useful functions take an atomic vector and return a list. For example, in [strings] you learned about `stringr::str_split()` which takes a character vector and returns a list of character vectors. If you use that inside mutate, you'll get a list-column:

```{r}
df <- tribble(
  ~x1,
  "a,b,c", 
  "d,e,f,g"
) 

df %>% 
  mutate(x2 = stringr::str_split(x1, ","))
```

`unnest()` knows how to handle these lists of vectors:

```{r}
df %>% 
  mutate(x2 = stringr::str_split(x1, ",")) %>% 
  unnest()
```

(If you find yourself using this pattern a lot, make sure to check out `tidyr::separate_rows()` which is a wrapper around this common pattern).

Another example of this pattern is using the `map()`, `map2()`, `pmap()` from purrr. For example, we could take the final example from [Invoking different functions] and rewrite it to use `mutate()`:

```{r}
sim <- tribble(
  ~f,      ~params,
  "runif", list(min = -1, max = 1),
  "rnorm", list(sd = 5),
  "rpois", list(lambda = 10)
)

sim %>%
  mutate(sims = invoke_map(f, params, n = 10))
```

Note that technically `sim` isn't homogeneous because it contains both double and integer vectors. However, this is unlikely to cause many problems since integers and doubles are both numeric vectors.

### From multivalued summaries

One restriction of `summarise()` is that it only works with summary functions that return a single value. That means that you can't use it with functions like `quantile()` that return a vector of arbitrary length:

```{r, error = TRUE}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = quantile(mpg))
```

You can however, wrap the result in a list! This obeys the contract of `summarise()`, because each summary is now a list (a vector) of length 1.

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(q = list(quantile(mpg)))
```

To make useful results with unnest, you'll also need to capture the probabilities:

```{r}
probs <- c(0.01, 0.25, 0.5, 0.75, 0.99)
mtcars %>% 
  group_by(cyl) %>% 
  summarise(p = list(probs), q = list(quantile(mpg, probs))) %>% 
  unnest()
```

### From a named list

Data frames with list-columns provide a solution to a common problem: what do you do if you want to iterate over both the contents of a list and its elements? Instead of trying to jam everything into one object, it's often easier to make a data frame: one column can contain the elements, and one column can contain the list.  An easy way to create such a data frame from a list is `tibble::enframe()`.  

```{r}
x <- list(
  a = 1:5,
  b = 3:4, 
  c = 5:6
) 

df <- enframe(x)
df
```

The advantage of this structure is that it generalises in a straightforward way - names are useful if you have character vector of metadata, but don't help if you have other types of data, or multiple vectors.

Now if you want to iterate over names and values in parallel, you can use `map2()`:

```{r}
df %>% 
  mutate(
    smry = map2_chr(name, value, ~ stringr::str_c(.x, ": ", .y[1]))
  )
```

### Exercises

1.  List all the functions that you can think of that take a atomic vector and 
    return a list.
    
1.  Brainstorm useful summary functions that, like `quantile()`, return
    multiple values.
    
1.  What's missing in the following data frame? How does `quantile()` return
    that missing piece? Why isn't that helpful here?

    ```{r}
    mtcars %>% 
      group_by(cyl) %>% 
      summarise(q = list(quantile(mpg))) %>% 
      unnest()
    ```

1.  What does this code do? Why might might it be useful?

    ```{r, eval = FALSE}
    mtcars %>% 
      group_by(cyl) %>% 
      summarise_each(funs(list))
    ```

## Simplifying list-columns

To apply the techniques of data manipulation and visualisation you've learned in this book, you'll need to simplify the list-column back to a regular column (an atomic vector), or set of columns. The technique you'll use to collapse back down to a simpler structure depends on whether you want a single value per element, or multiple values:

1.  If you want a single value, use `mutate()` with `map_lgl()`, 
    `map_int()`, `map_dbl()`, and `map_chr()` to create an atomic vector.
    
1.  If you want many values, use `unnest()` to convert list-columns back
    to regular columns, repeating the rows as many times as necessary.

These are described in more detail below.

### List to vector

If you can reduce your list column to an atomic vector then it will be a regular column. For example, you can always summarise an object with its type and length, so this code will work regardless of what sort of list-column you have:

```{r}
df <- tribble(
  ~x,
  letters[1:5],
  1:3,
  runif(5)
)
  
df %>% mutate(
  type = map_chr(x, typeof),
  length = map_int(x, length)
)
```

This is the same basic information that you get from the default tbl print method, but now you can use it for filtering. This is a useful technique if you have a heterogeneous list, and want to filter out the parts aren't working for you.

Don't forget about the `map_*()` shortcuts - you can use `map_chr(x, "apple")` to extract the string stored in `apple` for each element of `x`. This is useful for pulling apart nested lists into regular columns. Use the `.null` argument to provide a value to use if the element is missing (instead of returning `NULL`):

```{r}
df <- tribble(
  ~x,
  list(a = 1, b = 2),
  list(a = 2, c = 4)
)
df %>% mutate(
  a = map_dbl(x, "a"),
  b = map_dbl(x, "b", .null = NA_real_)
)
```

### Unnesting

`unnest()` works by repeating the regular columns once for each element of the list-column. For example, in the following very simple example we repeat the first row 4 times (because there the first element of `y` has length four), and the second row once:

```{r}
tibble(x = 1:2, y = list(1:4, 1)) %>% unnest(y)
```

This means that you can't simultaneously unnest two columns that contain different number of elements:

```{r, error = TRUE}
# Ok, because y and z have the same number of elements in
# every row
df1 <- tribble(
  ~x, ~y,           ~z,
   1, c("a", "b"), 1:2,
   2, "c",           3
)
df1
df1 %>% unnest(y, z)

# Doesn't work because y and z have different number of elements
df2 <- tribble(
  ~x, ~y,           ~z,
   1, "a",         1:2,  
   2, c("b", "c"),   3
)
df2
df2 %>% unnest(y, z)
```

The same principle applies when unnesting list-columns of data frames. You can unnest multiple list-cols as long as all the data frames in each row have the same number of rows.

### Exercises

1.  Why might the `lengths()` function be useful for creating atomic
    vector columns from list-columns?
    
1.  List the most common types of vector found in a data frame. What makes
    lists different?

## Making tidy data with broom

The broom package provides three general tools for turning models into tidy data frames:

1.  `broom::glance(model)` returns a row for each model. Each column gives a 
    model summary: either a measure of model quality, or complexity, or a 
    combination of the two.
   
1.  `broom::tidy(model)` returns a row for each coefficient in the model. Each 
    column gives information about the estimate or its variability.
    
1.  `broom::augment(model, data)` returns a row for each row in `data`, adding
    extra values like residuals, and influence statistics.

